---
title: "Join SDTM with its SUPP dataset (function)"
author: "Noory Kim"
date: "started 2025-05-19, updated 2025-05-22"
output: 
  pdf_document
  # pdf_document:
  #   latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE}
library(safetyData) # a package with sample CDISC data sets
library(tidyverse)
```

NOTES: 

- with tidyverse, column names are case sensitive

- tidyr
  - pivot_wider() : transpose from long to wide

- dplyr 
  - select() : keep/drop columns
    - select(-any_of(vars)) : drop columns
  - slice_head(n=) : keep the first n rows
  
- References: 
  - select(-any_of(vars)): https://tidyselect.r-lib.org/reference/all_of.html
  
UPDATES

- 2025-05-22: Added parameters ds_name and supp_name to the function

```{r datasets_in_safetyData}
## Get list of datasets
dataset_list <- data(package = "safetyData")$results[ , "Item"]
dataset_list
```

# The function

Parameters

- domain: SDTM domain abbreviation 

- ds_name: name of main dataset

- supp_name (default: NULL): name of existing SUPP-- dataset

```{r function}
join_domain_with_supp_by_seq <- function(domain, ds_name, supp_name=NULL){

## concatenate string to get name of datasets to join and of the SEQ variable
  name_main <- paste0(ds_name)
  
  if(!is.null(supp_name)){
    name_supp <- paste0(supp_name)
    name_seq  <- paste0(toupper(domain), "SEQ")
  } 
  
## get main dataset
  if(is.null(supp_name)){
    output <- get(name_main) %>% 
      as_tibble() 
  }
  else if(!is.null(supp_name)){
    main <- get(name_main) %>% 
      as_tibble() %>% 
      rename("SEQ"=name_seq) ## rename --SEQ as SEQ, to simplify join statement below
    
    ## Columns in SUPP-- not needed for ADaMs or TLFs 
    cols_to_drop <- c("STUDYID", "RDOMAIN", "IDVAR", "QLABEL", "QORIG", "QEVAL")
    
    ## get SUPP-- dataset and transpose
    supp_t <- get(name_supp) %>% 
      as_tibble() %>% 
      pivot_wider(names_from=QNAM, values_from=QVAL) %>% 
      mutate(SEQ = as.numeric(IDVARVAL)) %>% 
      select(-any_of(cols_to_drop))

    ## join datasets 
    output <- left_join(main, supp_t, by=c("USUBJID"="USUBJID", "SEQ"="SEQ")) %>% 
      rename(!!name_seq := "SEQ") ## rename SEQ back to --SEQ, now that the join is done
  }
  
  ## output result
  return(output)
}
```

# Function calls

## DS domain
```{r ds, message=FALSE}
ds <- join_domain_with_supp_by_seq(domain = "ds", ds_name = "sdtm_ds", supp_name = "sdtm_suppds")
```

Warning appears even though any_of() is being used.

Documented workaround shows an example for all_of(): 
https://tidyselect.r-lib.org/reference/faq-external-vector.html

```{r ds_tibble_slice_head}
# show the first few rows of a tibble
ds %>% 
  slice_head(n=3)
```

```{r ds_dataframe_head}
# show as a dataframe rather than as a tibble (which gets truncated)
ds %>% 
  as.data.frame() %>% 
  head(3)
```

## AE domain, with SUPPAE
```{r ae}
ae <- join_domain_with_supp_by_seq(domain = "ae", ds_name = "sdtm_ae", supp_name = "sdtm_suppae")

# show as a dataframe rather than as a tibble (which gets truncated)
ae %>% 
  as.data.frame() %>% 
  head(3)
```

## QS domain, without SUPP--
```{r qs}
qs <- join_domain_with_supp_by_seq(domain = "qs", ds_name = "sdtm_qs")

# show as a dataframe rather than as a tibble (which gets truncated)
qs %>% 
  as.data.frame() %>% 
  head(3)
```
